The Calibration Reference Data System

### 2015-10-06 docs/paper/crds.tex
added references
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2124 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3
various improvements, extra examples, section on selectors, etc
git-svn-id: https://aeon.stsci.edu/ssb/svn/crds_server/trunk@2128 37088bc5-bd3f-4c5b-b9de-119bb1e8f7e3

\author[stsci]{Perry~Greenfield}
\ead{perry@stsci.edu}
\author[stsci]{J.~Miller}

@@

FITS \sep
File formats \sep
Standards \sep
World Coordinate System

@@

@@

Figure~\ref{fig:example1} illustrates use of the first two items. Essentially two different configurations are supported, and for each there are three possible choices of reference files that depend on the date of the observation. For example, if the date of the observation for the configuration {\tt (\textquotesingle MIRIFULONG\textquotesingle , \textquotesingle FAST\textquotesingle , \textquotesingle FULL\textquotesingle )} is 2015-04-02, the second choice will be used resulting in the return of {\tt jwst\_miri\_dark\_0032.fits} as the recommended reference file.

\begin{figure*}
\begin{verbatim}
selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

Figure~\ref{fig:example2} shows the use of the software selection mechanism. In this case if the software version is less than 3.1 (using the typical version semantics) then the first file will be chosen, if less than 5 the second will be chosen, otherwise the last file will be used.

@@

[xxx need headers included and explained since examples have unnamed
parameters that won't make any sense unless put in context]

[xxx insert full description of syntax used]

[xxx somewhere put in explanation of safety features preventing hand
mangling of rules files and configuration]

Full documentation of the syntax and semantics of all the rmap and
context files is available on the web 
\footnote{\texttt{https://jwst-crds.stsci.edu/static/users\_guide/rmap\_syntax.html\#selectors}}

@@

[xxx need example of instrument context file]

@@

[xxx need example of pipeline context file]

@@

@@

@@

\bibliographystyle{model2-names}
\bibliography{asdf}

>>>>

\author[stsci]{Perry~Greenfield}
\ead{perry@stsci.edu}
\author[stsci]{Todd~Miller}

@@

Calibration \sep
Pipelines \sep

@@

\subsection{Rules Framework}

All the files that define the rules for how reference files are selected
use the same organization and syntax.
A simple syntax is used that consists of expected variables set to simple, 
possibly nested, datastructures. Each file is expected to have a header 
and selector defined. The header contains information about the intended 
use of the file, history, information to validate the integrity of the file,
and other similar items. Most important, the header identifies the exposure
parameters that are to be used in the selection process. The selector 
contains the specifics on the rules themselves.

@@

Figure~\ref{fig:example1} illustrates use of the first two items. First note the contents
of the header. It shows what previous rmap file it was derived from, that it is defining
the rules for the selection of reference files (since this syntax is also used for
the selection of other kinds of files), what kind of reference file is being selected, information about the telescope and instrument, and what the name of the file is. 
The {\tt parkey} attribute defines the parameters to use for the selection. In the
case of JWST, the parameters map to FITS header keywords, but are referred through
nested object-oriented scheme; the mapping to specific keywords is handled by a special
library interface that is part of the calibration pipeline. In this specific example,
This is the rules file for selecting the JWST MIRI instrument dark files, based on 
the specific detector, read pattern, and subarray configuration, along with the 
observation date.

The header permits attributes that allow specialized handling of the metadata 
the dataset, including indicating when parameters are relevant or not (e.g., only 
care about its value if the value of another parameter is given as specified), 
how to handle cases where no match is found, and hooks to allow special preprocessing
of parameter values. See the documentation for details on these cases. 
\footnote{\url{https://jwst-crds.stsci.edu/static/users\_guide/rmap_syntax.html\#active-header-fields}}
Most of these special features were added to handle the existing complexitites present in the CDBS rule system; we expect these will be used far less for JWST.

The selector part of the file indicates that a match is first performed on the 
three specified parameters. There are two supported configurations in this example.
If one matches, then the date is used to determine which of the following three files
should be used.
For example, if the date of the observation for the configuration {\tt (\textquotesingle 
MIRIFULONG\textquotesingle , \textquotesingle FAST\textquotesingle , 
\textquotesingle FULL\textquotesingle )} is 2015-04-02, the second choice 
will be used resulting in the return of {\tt jwst\_miri\_dark\_0032.fits} 
as the recommended dark reference file.

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_miri_dark_0007.rmap',
    'filekind' : 'DARK',
    'instrument' : 'MIRI',
    'mapping' : 'REFERENCE',
    'name' : 'jwst_miri_dark_0009.rmap',
    'observatory' : 'JWST',
    'parkey' : (('META.INSTRUMENT.DETECTOR', 'META.EXPOSURE.READPATT', 'META.SUBARRAY.NAME'),),
    'sha1sum' : 'bd99f4960096de9ad39a9ce506d5ff985847c4a3',
}

selector = Match({
    ('MIRIFULONG', 'FAST', 'FULL') : UseAfter({
        '2015-01-27 12:05:34' 'jwst_miri_dark_0029.fits',
        '2015-03-22 00:29:01' 'jwst_miri_dark_0032.fits',
        '2015-04-07 09:14:00' 'jwst_miri_dark_0047.fits',
        ...
        }),
    ('MIRIFUSHORT', 'SLOW', 'FULL') : UseAfter({
        '2015-01-25 12:06:34' 'jwst_miri_dark_0099.fits',
        '2015-03-21 00:28:01' 'jwst_miri_dark_0101.fits',
        '2015-04-08 09:13:00' 'jwst_miri_dark_0127.fits',
        ...
        }),
        ...
})
\end{verbatim}
\caption{Example showing basic parameter matching rules coupled with date dependency.}
\label{fig:example1}
\end{figure*}

Figure~\ref{fig:example2} shows the use of the software selection mechanism (for brevity, the header is omitted in this and subsequent examples). In this case if the software version is less than 3.1 (using the typical version semantics) then the first file will be chosen, if less than 5 the second will be chosen, otherwise the last file will be used.

@@

\subsection{Suppported Selectors and Features}

This section more systematically describes existing selector mechanisms. 
First is is necessary to understand how the supplied parameters (specified
in the header parkey attribute) are used with selectors. If selectors are 
nested, each layer removes one or more parameters from the list leaving 
the rest for lower nestings. For example, if parkey is as specified in
Figure~\ref{fig:example1}, then 4 parameters are available. The Match 
selector is requesting 3 since all of its keys have a tuple of 3 items,
that leaves only the date parameter for subsequent selectors.
In all cases, the selector expects to be provided an argument that is the 
appropriate type (e.g., if time).

Selectors may return a single filename, a tuple of filenames,
``Not Applicable'' (which basically implies that no file is
needed for this specific case, i.e., no such calibration need be 
performed) or ``OMIT'' (Currently unused).

\subsubsection{Match}
This selector finds which tuple of parameters matches the supplied tuple
of values for those parameters. This is a very flexible selector, and 
many options are provided for its use. (If only one parameteter is being,
matched, the tuple can be replaced by a simple string.)

In the simple case, matches are required to be exact. However, to avoid
rules from having to specify large numbers of combinations where no
distinctions are being made, a number of mechanisms have been provided
to make rules much more concise.

\begin{description}
\item[lists of allowable alternatives:] Allowable values for parameteters
may be supplied as an \textquotesingle OR\textquotesingle 
list using {\tt |} as a separator. For example:
\texttt {({\ttfamily"}filter1{\ttfamily"}
| {\ttfamily"}filter2{\ttfamily"},...)} will allow either value to match the first
parameter.
\item[Wild Cards:] Using an {\tt \*} in a string applies familiar wildcard 
matching rules to the supplied string. For example 
\texttt {({\ttfamily"}F*122{\ttfamily"})} will
match any value starting with {\tt F} and ending with {\tt 122}
\item[Regular Expressions:] This allows matching true regular expressions,
allowing for more complex matching rules than simple wildcards permit.
\item[Literal Expressions:] A literal expression is bracketed with curly
braces {\tt \{ \}} and allows inclusion of special characters without 
special interpretation by CRDS, such as {\tt | and \*}.
\item[Relational Expressions:] Relational expressions are braketed by {\tt \#}.
These do numerical comparisons on the supplied parameter value to determine
a match. Supported comparison operators are: 
\texttt {\textgreater, \textgreater =, \textless, \textless =, == and or}. 
For example \texttt {({\ttfamily"}\# >=1 and <37\",)} 
will match any number equal to or 
greater than 1 and less than 37.
\item[Between:] This is simply a more convenient way of expressing a range
than using relational expressions. E.g., equivalent to the previous example
is: \texttt{{\ttfamily"}between 1 37{\ttfamily"}}. It is inclusive of the first number.
\item[N/A:] This is equivalent to using a wildcard for the value, but with 
the difference of how weighting is used (discussed below). This does not
affect the weight of the match.
\item[NOT Expressions:] A negation of a match. 
\texttt {{\ttfamily"}not HRC{\ttfamily"}} 
is a match if the value does not match {\tt HRC}.
\item[Substitutions:] A shorthand notation that needs better explanation 
than is in the documenation [I couldn't understatnd it as written xxx]
\end{description}

With all these match options it is possible to write multiple items that 
match a given case. In order to allow effective defaults for general cases
while allowing specific cases to override the general case, special rule
for disambiguation in these situations have been developed. The basic idea
is that more specific matches take precedence over more general matches.
For example, if one item lists as its matching value a wildcard, and another
gives a specific value for the same parameter, the specific match takes
precedence. The precedence is handled by determining the generality of the
match. Use of wildcards or regular expressions increase the generality index,
while use of N/A does not. [I found the explanation of Match weighting
unclear, and it suggests the oppposite, a couple examples would help a lot
 xxx].

If after using these precedence rules, there are still multiple matches,
an error results.

\subsubsection{UseAfter}
This expects an ordered sequence of date/time values as part of the rules.
This selector finds the greates date/time value that is less than or equal
to the supplied date/time.

\subsubsection{SelectVersion}
This is essentially a special case of relational matches that understands
the semantics of software versions (e.g., that 1.2 > 1.1.10).

\subsubsection{ClosestTime}
This selector selects the item with the closest time.

\subsubsection{GeometricallyNearest}
This selector selects the item where the ratio of the parameter to that
in the list of choices is closest to 1. For example, 4 would match 10 
instead of 1 since 10/4 is closer to 1 than 4/1.

\subsubsection{Bracket}
This selector finds the two items for which one is next greater in value,
and the one that is next least in value (as illustrated in
Figure~\ref{fig:example3}).

Full documentation of the syntax and semantics of all the rmap and
context files is available on the web 
\footnote{\url{https://jwst-crds.stsci.edu/static/users\_guide/rmap\_syntax.html\#selectors}}

@@

Figure~\ref{fig:example_instrument_context}

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_nircam_0015.imap',
    'instrument' : 'NIRCAM',
    'mapping' : 'INSTRUMENT',
    'name' : 'jwst_nircam_0016.imap',
    'observatory' : 'JWST',
    'parkey' : ('REFTYPE',),
    'sha1sum' : '0fd29b57f226f8c53bf3dfa7564eccd4e3fc56bf',
}

selector = {
    'DARK' : 'jwst_nircam_dark_0003.rmap',
    'DISTORTION' : 'N/A',
    'DRIZPARS' : 'jwst_nircam_drizpars_0000.rmap',
    'FLAT' : 'jwst_nircam_flat_0000.rmap',
    'GAIN' : 'jwst_nircam_gain_0001.rmap',
    'IPC' : 'jwst_nircam_ipc_0000.rmap',
    'LASTFRAME' : 'N/A',
    'LINEARITY' : 'jwst_nircam_linearity_0003.rmap',
    'MASK' : 'jwst_nircam_mask_0002.rmap',
    'PHOTOM' : 'jwst_nircam_photom_0002.rmap',
    'READNOISE' : 'jwst_nircam_readnoise_0001.rmap',
    'REGIONS' : 'N/A',
    'RESET' : 'N/A',
    'SATURATION' : 'jwst_nircam_saturation_0002.rmap',
    'SPECWCS' : 'N/A',
    'SUPERBIAS' : 'jwst_nircam_superbias_0003.rmap',
    'WCSREGIONS' : 'N/A',
}
\end{verbatim}
\caption{An example of an instrument contex file (.imap)}
\label{fig:example_instrument_context}
\end{figure*}

@@

Figure~\ref{fig:example_pipeline_context} shows an example of such a pipeline context file.
It is used to select the appropriate instrument context file to use.

\begin{figure*}
\begin{verbatim}
header = {
    'derived_from' : 'jwst_0068.pmap',
    'description' : 'Hand edited to support updated .imaps for N/A',
    'mapping' : 'PIPELINE',
    'name' : 'jwst_0069.pmap',
    'observatory' : 'JWST',
    'parkey' : ('META.INSTRUMENT.NAME',),
    'sha1sum' : '691d78053ef710f31bfd8f5bea8ffdc3c1dec1aa',
}

selector = {
    'FGS' : 'jwst_fgs_0016.imap',
    'MIRI' : 'jwst_miri_0043.imap',
    'NIRCAM' : 'jwst_nircam_0016.imap',
    'NIRISS' : 'jwst_niriss_0024.imap',
    'NIRSPEC' : 'jwst_nirspec_0026.imap',
}
\end{verbatim}
\caption{An example of a pipeline context file (.pmap)}
\label{fig:example_pipeline_context}
\end{figure*}

@@

\subsection{Ensuring integrity of the configuration}

Since all the rules involve simple text files, one possible corruption of the
integrity of the system is if someone edits or otherwise changes the
contents of one of the files. As a protection against that, all the files
involved contain a SHA1SUM checksum on the contents 
(\citealt{sha1ref}).
CRDS when loading the
files compares the the computed SHA1SUM against what is contained in the 
file header. If they do not match, an error is raised.

To allow flexibility in usage by remote users of the pipeline, one can
invoke CRDS so that this check is overridden, allowing users to hand edit 
such files with their local customizations. But with the operational systems,
these checks are always performed.

@@

Another improvement planned is the ability to return values other than
filenames or tuples of filenames. A common use of reference files for HST 
was to provide a table of parameter to be used by the calibration 
pipeline to tune processing to the configurations. Often the real need
of the pipeline was just a handful of parameter values. Often this resulted
in a dual layer lookup system. CDBS would select a table (though sometimes
there was only one reference file for all cases), and then the calibration
pipeline would select a specific row in the table that matched an additional
set of selection parameters. This could be greately simplified if the 
rmap itself bypassed reference files entirely and effectively embedded
the parameter values as what is returned. In this manner, it relieves 
the calibration pipeline of embedding its own matching system in its code.

@@

\bibliographystyle{model2-names}
\bibliography{crds}
